---
title: "Tloop_project_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(RColorBrewer)
library(plyr)
library(dplyr)
library(tibble)
library(ggsignif)
library(plotrix)
library(openxlsx)
```

## Creating the function to remove outliers based on Interquartile range
```{r}
removeOut <- function(distribution, bounds=1.5, na.rm = TRUE) {
 Q <- quantile(distribution, probs=c(.25,.75), na.rm=TRUE) 
 #finds the 25th and 75th percentile of the distribution, excluding NAs
 iqr <- IQR(distribution, na.rm=TRUE)
 #difference between 75th and 25th quartile=interquartile range
 upperBound <- Q[2]+bounds*iqr
 #upper bound of the values is 75th quartile + 1.5*(interquartile range) note: that 1.5 can be vaired by changing bound
 lowerBound <- Q[1]-bounds*iqr
 #upper bound of the values is 25th quartile - 1.5*(interquartile range)
 eliminated <- c(distribution < lowerBound | distribution > upperBound)
 #take out the values that are outside of the established bounds note: that 1.5 can be vaired by changing bound
 noteliminatedindex <- which(distribution > lowerBound & distribution < upperBound)
 #index values within the upper and lower bounds
 noteliminatedvalues <- distribution[noteliminatedindex]
 return(noteliminatedvalues)
 #return those values which will be included that are between the upper and lower bounds 
}
```

Read in csv.
Formatting of csv: each column is a different colony, first row is the name of the colony, each column is then filled with the background fluorescence corrected flourescence values (there should be three values per biological replicate, each of which are the technical replicates). Far right column is the condition of the growth, either - or + ligand. For this particular switch, -L=repression and +L=expression. For those colonies that have less raw data values than the controls, the empty cells are filled with an integer that can be removed easily (-2) to be turned into NA when inputted into R. 
Put NA in place of -2 (which was how NA was denoted in the csv).
This is for repression and expression, fold induction will not be included because it will be calculated based on the included repression and expression values.
```{r}
#LIBRARY 1
#read in csv, these values will include colonies below the function cut off which has been defined here as 2.0
TloopL1all <- read.csv("TloopL1.csv", header=T)
empty_columns <- sapply(TloopL1all, function(x) all(is.na(x) | x==""))
TloopL1all <- TloopL1all[,!empty_columns]
TloopL1all[TloopL1all == -2] <-NA
FIrows <- which(TloopL1all$COLONIES == "FI")

#MINUS CNCbl DF
TloopL1_minus1 <- TloopL1all[TloopL1all == "(-)",]
TloopL1_minus <- TloopL1_minus1[rowSums(is.na(TloopL1_minus1)) !=ncol(TloopL1_minus1),]
rm(TloopL1_minus1)

#PLUS CNCbl DF
TloopL1_plus1 <- TloopL1all[TloopL1all == "(+)",]
TloopL1_plus <- TloopL1_plus1[rowSums(is.na(TloopL1_plus1)) !=ncol(TloopL1_plus1),]
rm(TloopL1_plus1)

#FI DF
TloopL1_FI1 <- TloopL1all[TloopL1all == "FI",]
#dataframe of just fold induction
TloopL1_FI <- TloopL1_FI1[rowSums(is.na(TloopL1_FI1)) !=ncol(TloopL1_FI1),]
rm(TloopL1_FI1)


#LIBRARY 2
#read in csv, these values will include colonies below the function cut off which has been defined here as 2.0
TloopL2all <- read.csv("TloopL2.csv", header=T)
empty_columns <- sapply(TloopL2all, function(x) all(is.na(x) | x==""))
TloopL2all <- TloopL2all[,!empty_columns]
TloopL2all[TloopL2all == -2] <-NA
FIrows <- which(TloopL2all$COLONIES == "FI")

#MINUS CNCbl DF
TloopL2_minus1 <- TloopL2all[TloopL2all == "(-)",]
TloopL2_minus <- TloopL2_minus1[rowSums(is.na(TloopL2_minus1)) !=ncol(TloopL2_minus1),]
rm(TloopL2_minus1)

#PLUS CNCbl DF
TloopL2_plus1 <- TloopL2all[TloopL2all == "(+)",]
TloopL2_plus <- TloopL2_plus1[rowSums(is.na(TloopL2_plus1)) !=ncol(TloopL2_plus1),]
rm(TloopL2_plus1)

#FI DF
TloopL2_FI1 <- TloopL2all[TloopL2all == "FI",]
#dataframe of just fold induction
TloopL2_FI <- TloopL2_FI1[rowSums(is.na(TloopL2_FI1)) !=ncol(TloopL2_FI1),]
rm(TloopL2_FI1)


#LIBRARY 3
#read in csv, these values will include colonies below the function cut off which has been defined here as 2.0
TloopL3all <- read.csv("TloopL3.csv", header=T)
empty_columns <- sapply(TloopL3all, function(x) all(is.na(x) | x==""))
TloopL3all <- TloopL3all[,!empty_columns]
TloopL3all[TloopL3all == -2] <-NA
FIrows <- which(TloopL3all$COLONIES == "FI")

#MINUS CNCbl DF
TloopL3_minus1 <- TloopL3all[TloopL3all == "(-)",]
TloopL3_minus <- TloopL3_minus1[rowSums(is.na(TloopL3_minus1)) !=ncol(TloopL3_minus1),]
rm(TloopL3_minus1)


#PLUS CNCbl DF
TloopL3_plus1 <- TloopL3all[TloopL3all == "(+)",]
TloopL3_plus <- TloopL3_plus1[rowSums(is.na(TloopL3_plus1)) !=ncol(TloopL3_plus1),]
rm(TloopL3_plus1)

#FI DF
TloopL3_FI1 <- TloopL3all[TloopL3all == "FI",]
#dataframe of just fold induction
TloopL3_FI <- TloopL3_FI1[rowSums(is.na(TloopL3_FI1)) !=ncol(TloopL3_FI1),]
rm(TloopL3_FI1)

#LIBRARY 4
#read in csv, these values will include colonies below the function cut off which has been defined here as 2.0
TloopL4all <- read.csv("TloopL4.csv", header=T)
empty_columns <- sapply(TloopL4all, function(x) all(is.na(x) | x==""))
TloopL4all <- TloopL4all[,!empty_columns]
TloopL4all[TloopL4all == -2] <-NA
FIrows <- which(TloopL4all$COLONIES == "FI")
#MINUS CNCbl DF
TloopL4_minus1 <- TloopL4all[TloopL4all == "(-)",]
TloopL4_minus <- TloopL4_minus1[rowSums(is.na(TloopL4_minus1)) !=ncol(TloopL4_minus1),]
rm(TloopL4_minus1)

#PLUS CNCbl DF
TloopL4_plus1 <- TloopL4all[TloopL4all == "(+)",]
TloopL4_plus <- TloopL4_plus1[rowSums(is.na(TloopL4_plus1)) !=ncol(TloopL4_plus1),]
rm(TloopL4_plus1)

#FI DF
TloopL4_FI1 <- TloopL4all[TloopL4all == "FI",]
#dataframe of just fold induction
TloopL4_FI <- TloopL4_FI1[rowSums(is.na(TloopL4_FI1)) !=ncol(TloopL4_FI1),]
rm(TloopL4_FI1)

#LIBRARY 5
#read in csv, these values will include colonies below the function cut off which has been defined here as 2.0
TloopL5all <- read.csv("TloopL5.csv", header=T)
empty_columns <- sapply(TloopL5all, function(x) all(is.na(x) | x==""))
TloopL5all <- TloopL5all[,!empty_columns]
TloopL5all[TloopL5all == -2] <-NA
FIrows <- which(TloopL5all$COLONIES == "FI")

#MINUS CNCbl DF
TloopL5_minus1 <- TloopL5all[TloopL5all == "(-)",]
TloopL5_minus <- TloopL5_minus1[rowSums(is.na(TloopL5_minus1)) !=ncol(TloopL5_minus1),]
rm(TloopL5_minus1)

#PLUS CNCbl DF
TloopL5_plus1 <- TloopL5all[TloopL5all == "(+)",]
TloopL5_plus <- TloopL5_plus1[rowSums(is.na(TloopL5_plus1)) !=ncol(TloopL5_plus1),]
rm(TloopL5_plus1)
#FI DF
TloopL5_FI1 <- TloopL5all[TloopL5all == "FI",]
#dataframe of just fold induction
TloopL5_FI <- TloopL5_FI1[rowSums(is.na(TloopL5_FI1)) !=ncol(TloopL5_FI1),]
rm(TloopL5_FI1)

#TOP SWITCHERS
#read in csv, these values will include colonies below the function cut off which has been defined here as 2.0
TloopTop10all <- read.csv("Tloop_top10_data.csv", header=T)
empty_columns <- sapply(TloopTop10all, function(x) all(is.na(x) | x==""))
TloopTop10all <- TloopTop10all[,!empty_columns]
TloopTop10all[TloopTop10all == -2] <-NA
FIrows <- which(TloopTop10all$COLONIES == "FI")

#MINUS CNCbl DF
TloopTop10_minus1 <- TloopTop10all[TloopTop10all == "(-)",]
TloopTop10_minus <- TloopTop10_minus1[rowSums(is.na(TloopTop10_minus1)) !=ncol(TloopTop10_minus1),]
rm(TloopTop10_minus1)


#PLUS CNCbl DF
TloopTop10_plus1 <- TloopTop10all[TloopTop10all == "(+)",]
TloopTop10_plus <- TloopTop10_plus1[rowSums(is.na(TloopTop10_plus1)) !=ncol(TloopTop10_plus1),]
rm(TloopTop10_plus1)
#FI DF
TloopTop10_FI1 <- TloopTop10all[TloopTop10all == "FI",]
#dataframe of just fold induction
TloopTop10_FI <- TloopTop10_FI1[rowSums(is.na(TloopTop10_FI1)) !=ncol(TloopTop10_FI1),]
rm(TloopTop10_FI1)
```
###Using function to remove outliers and create a new dataframe in wide form and long form for minus
```{r}
##LIBRARY 1
noOut_TloopL1_minus.list <- apply(TloopL1_minus[,2:ncol(TloopL1_minus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL1_minus_sideways <- plyr::ldply(noOut_TloopL1_minus.list, rbind)
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL1_minus <- data.frame(t(noOut_TloopL1_minus_sideways[-1]))
colnames(noOut_TloopL1_minus) <- noOut_TloopL1_minus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL1_minus_noOut_wideform <- cbind(TloopL1_minus[1:nrow(noOut_TloopL1_minus),1], noOut_TloopL1_minus[,1:ncol(noOut_TloopL1_minus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL1_minus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL1_minus <- gather(TloopL1_minus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
##LIBRARY 2
noOut_TloopL2_minus.list <- apply(TloopL2_minus[,2:ncol(TloopL2_minus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL2_minus_sideways <- plyr::ldply(noOut_TloopL2_minus.list, rbind)
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL2_minus <- data.frame(t(noOut_TloopL2_minus_sideways[-1]))
colnames(noOut_TloopL2_minus) <- noOut_TloopL2_minus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL2_minus_noOut_wideform <- cbind(TloopL2_minus[1:nrow(noOut_TloopL2_minus),1], noOut_TloopL2_minus[,1:ncol(noOut_TloopL2_minus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL2_minus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL2_minus <- gather(TloopL2_minus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
##LIBRARY 3
noOut_TloopL3_minus.list <- apply(TloopL3_minus[,2:ncol(TloopL3_minus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL3_minus_sideways <- plyr::ldply(noOut_TloopL3_minus.list, rbind)
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL3_minus <- data.frame(t(noOut_TloopL3_minus_sideways[-1]))
colnames(noOut_TloopL3_minus) <- noOut_TloopL3_minus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL3_minus_noOut_wideform <- cbind(TloopL3_minus[1:nrow(noOut_TloopL3_minus),1], noOut_TloopL3_minus[,1:ncol(noOut_TloopL3_minus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL3_minus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL3_minus <- gather(TloopL3_minus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
##LIBRARY 4
noOut_TloopL4_minus.list <- apply(TloopL4_minus[,2:ncol(TloopL4_minus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL4_minus_sideways <- plyr::ldply(noOut_TloopL4_minus.list, rbind)
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL4_minus <- data.frame(t(noOut_TloopL4_minus_sideways[-1]))
colnames(noOut_TloopL4_minus) <- noOut_TloopL4_minus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL4_minus_noOut_wideform <- cbind(TloopL4_minus[1:nrow(noOut_TloopL4_minus),1], noOut_TloopL4_minus[,1:ncol(noOut_TloopL4_minus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL4_minus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL4_minus <- gather(TloopL4_minus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
##LIBRARY 5 
noOut_TloopL5_minus.list <- apply(TloopL5_minus[,2:ncol(TloopL5_minus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL5_minus_sideways <- plyr::ldply(noOut_TloopL5_minus.list, rbind)
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL5_minus <- data.frame(t(noOut_TloopL5_minus_sideways[-1]))
colnames(noOut_TloopL5_minus) <- noOut_TloopL5_minus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL5_minus_noOut_wideform <- cbind(TloopL5_minus[1:nrow(noOut_TloopL5_minus),1], noOut_TloopL5_minus[,1:ncol(noOut_TloopL5_minus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL5_minus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL5_minus <- gather(TloopL5_minus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
##TOP SWITCHES
noOut_TloopTop10_minus.list <- apply(TloopTop10_minus[,2:ncol(TloopTop10_minus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopTop10_minus_sideways <- plyr::ldply(noOut_TloopTop10_minus.list, rbind)
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopTop10_minus <- data.frame(t(noOut_TloopTop10_minus_sideways[-1]))
colnames(noOut_TloopTop10_minus) <- noOut_TloopTop10_minus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopTop10_minus_noOut_wideform <- cbind(TloopTop10_minus[1:nrow(noOut_TloopTop10_minus),1], noOut_TloopTop10_minus[,1:ncol(noOut_TloopTop10_minus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopTop10_minus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopTop10_minus <- gather(TloopTop10_minus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
```

###Using function to remove outliers and create a new dataframe in wide form and long form for plus
```{r}
##LIBRARY 1
noOut_TloopL1_plus.list <- apply(TloopL1_plus[,2:ncol(TloopL1_plus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL1_plus_sideways <- plyr::ldply(noOut_TloopL1_plus.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL1_plus <- data.frame(t(noOut_TloopL1_plus_sideways[-1]))
colnames(noOut_TloopL1_plus) <- noOut_TloopL1_plus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL1_plus_noOut_wideform <- cbind(TloopL1_plus[1:nrow(noOut_TloopL1_plus),1], noOut_TloopL1_plus[,1:ncol(noOut_TloopL1_plus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL1_plus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL1_plus <- gather(TloopL1_plus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
##LIBRARY 2
noOut_TloopL2_plus.list <- apply(TloopL2_plus[,2:ncol(TloopL2_plus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL2_plus_sideways <- plyr::ldply(noOut_TloopL2_plus.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL2_plus <- data.frame(t(noOut_TloopL2_plus_sideways[-1]))
colnames(noOut_TloopL2_plus) <- noOut_TloopL2_plus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL2_plus_noOut_wideform <- cbind(TloopL2_plus[1:nrow(noOut_TloopL2_plus),1], noOut_TloopL2_plus[,1:ncol(noOut_TloopL2_plus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL2_plus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL2_plus <- gather(TloopL2_plus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
##LIBRARY 3
noOut_TloopL3_plus.list <- apply(TloopL3_plus[,2:ncol(TloopL3_plus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL3_plus_sideways <- plyr::ldply(noOut_TloopL3_plus.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL3_plus <- data.frame(t(noOut_TloopL3_plus_sideways[-1]))
colnames(noOut_TloopL3_plus) <- noOut_TloopL3_plus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL3_plus_noOut_wideform <- cbind(TloopL3_plus[1:nrow(noOut_TloopL3_plus),1], noOut_TloopL3_plus[,1:ncol(noOut_TloopL3_plus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL3_plus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL3_plus <- gather(TloopL3_plus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
##LIBRARY 4
noOut_TloopL4_plus.list <- apply(TloopL4_plus[,2:ncol(TloopL4_plus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL4_plus_sideways <- plyr::ldply(noOut_TloopL4_plus.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL4_plus <- data.frame(t(noOut_TloopL4_plus_sideways[-1]))
colnames(noOut_TloopL4_plus) <- noOut_TloopL4_plus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL4_plus_noOut_wideform <- cbind(TloopL4_plus[1:nrow(noOut_TloopL4_plus),1], noOut_TloopL4_plus[,1:ncol(noOut_TloopL4_plus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL4_plus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL4_plus <- gather(TloopL4_plus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
##LIBRARY 5
noOut_TloopL5_plus.list <- apply(TloopL5_plus[,2:ncol(TloopL5_plus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL5_plus_sideways <- plyr::ldply(noOut_TloopL5_plus.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL5_plus <- data.frame(t(noOut_TloopL5_plus_sideways[-1]))
colnames(noOut_TloopL5_plus) <- noOut_TloopL5_plus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL5_plus_noOut_wideform <- cbind(TloopL5_plus[1:nrow(noOut_TloopL5_plus),1], noOut_TloopL5_plus[,1:ncol(noOut_TloopL5_plus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL5_plus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL5_plus <- gather(TloopL5_plus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
##TOP SWITCHES
noOut_TloopTop10_plus.list <- apply(TloopTop10_plus[,2:ncol(TloopTop10_plus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopTop10_plus_sideways <- plyr::ldply(noOut_TloopTop10_plus.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopTop10_plus <- data.frame(t(noOut_TloopTop10_plus_sideways[-1]))
colnames(noOut_TloopTop10_plus) <- noOut_TloopTop10_plus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopTop10_plus_noOut_wideform <- cbind(TloopTop10_plus[1:nrow(noOut_TloopTop10_plus),1], noOut_TloopTop10_plus[,1:ncol(noOut_TloopTop10_plus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopTop10_plus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopTop10_plus <- gather(TloopTop10_plus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)

```

###Using function to remove outliers and create a new dataframe in wide form and long form for FI
```{r}
##LIBRARY 1
noOut_TloopL1_FI.list <- apply(TloopL1_FI[,2:ncol(TloopL1_FI)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL1_FI_sideways <- plyr::ldply(noOut_TloopL1_FI.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL1_FI <- data.frame(t(noOut_TloopL1_FI_sideways[-1]))
colnames(noOut_TloopL1_FI) <- noOut_TloopL1_FI_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL1_FI_noOut_wideform <- cbind(TloopL1_FI[1:nrow(noOut_TloopL1_FI),1], noOut_TloopL1_FI[,1:ncol(noOut_TloopL1_FI)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL1_FI_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL1_FI <- gather(TloopL1_FI_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
medianL1_FI<- as.data.frame(apply(noOut_TloopL1_FI, 2, median, na.rm=TRUE))
##LIBRARY 2
noOut_TloopL2_FI.list <- apply(TloopL2_FI[,2:ncol(TloopL2_FI)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL2_FI_sideways <- plyr::ldply(noOut_TloopL2_FI.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL2_FI <- data.frame(t(noOut_TloopL2_FI_sideways[-1]))
colnames(noOut_TloopL2_FI) <- noOut_TloopL2_FI_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL2_FI_noOut_wideform <- cbind(TloopL2_FI[1:nrow(noOut_TloopL2_FI),1], noOut_TloopL2_FI[,1:ncol(noOut_TloopL2_FI)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL2_FI_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL2_FI <- gather(TloopL2_FI_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
medianL2_FI<- as.data.frame(apply(noOut_TloopL2_FI, 2, median, na.rm=TRUE))
##LIBRARY 3
noOut_TloopL3_FI.list <- apply(TloopL3_FI[,2:ncol(TloopL3_FI)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL3_FI_sideways <- plyr::ldply(noOut_TloopL3_FI.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL3_FI <- data.frame(t(noOut_TloopL3_FI_sideways[-1]))
colnames(noOut_TloopL3_FI) <- noOut_TloopL3_FI_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL3_FI_noOut_wideform <- cbind(TloopL3_FI[1:nrow(noOut_TloopL3_FI),1], noOut_TloopL3_FI[,1:ncol(noOut_TloopL3_FI)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL3_FI_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL3_FI <- gather(TloopL3_FI_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
medianL3_FI<- as.data.frame(apply(noOut_TloopL3_FI, 2, median, na.rm=TRUE))
##LIBRARY 4
noOut_TloopL4_FI.list <- apply(TloopL4_FI[,2:ncol(TloopL4_FI)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL4_FI_sideways <- plyr::ldply(noOut_TloopL4_FI.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL4_FI <- data.frame(t(noOut_TloopL4_FI_sideways[-1]))
colnames(noOut_TloopL4_FI) <- noOut_TloopL4_FI_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL4_FI_noOut_wideform <- cbind(TloopL4_FI[1:nrow(noOut_TloopL4_FI),1], noOut_TloopL4_FI[,1:ncol(noOut_TloopL4_FI)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL4_FI_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL4_FI <- gather(TloopL4_FI_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
medianL4_FI<- as.data.frame(apply(noOut_TloopL4_FI, 2, median, na.rm=TRUE))
##LIBRARY 5
noOut_TloopL5_FI.list <- apply(TloopL5_FI[,2:ncol(TloopL5_FI)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopL5_FI_sideways <- plyr::ldply(noOut_TloopL5_FI.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopL5_FI <- data.frame(t(noOut_TloopL5_FI_sideways[-1]))
colnames(noOut_TloopL5_FI) <- noOut_TloopL5_FI_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopL5_FI_noOut_wideform <- cbind(TloopL5_FI[1:nrow(noOut_TloopL5_FI),1], noOut_TloopL5_FI[,1:ncol(noOut_TloopL5_FI)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopL5_FI_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopL5_FI <- gather(TloopL5_FI_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
medianL5_FI<- as.data.frame(apply(noOut_TloopL5_FI, 2, median, na.rm=TRUE))
##TOP SWITCHES
noOut_TloopTop10_FI.list <- apply(TloopTop10_FI[,2:ncol(TloopTop10_FI)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_TloopTop10_FI_sideways <- plyr::ldply(noOut_TloopTop10_FI.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_TloopTop10_FI <- data.frame(t(noOut_TloopTop10_FI_sideways[-1]))
colnames(noOut_TloopTop10_FI) <- noOut_TloopTop10_FI_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
TloopTop10_FI_noOut_wideform <- cbind(TloopTop10_FI[1:nrow(noOut_TloopTop10_FI),1], noOut_TloopTop10_FI[,1:ncol(noOut_TloopTop10_FI)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(TloopTop10_FI_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_TloopTop10_FI <- gather(TloopTop10_FI_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
medianTop10_FI<- as.data.frame(apply(noOut_TloopTop10_FI, 2, median, na.rm=TRUE))
```
##Make DF of medians of each after removal of outliers 
```{r}
##LIBRARY 1
#median DFs
medianL1_FI<- as.data.frame(apply(noOut_TloopL1_FI, 2, median, na.rm=TRUE))
medianL1_minus<- as.data.frame(apply(noOut_TloopL1_minus, 2, median, na.rm=TRUE))
medianL1_plus<- as.data.frame(apply(noOut_TloopL1_plus, 2, median, na.rm=TRUE))
#standard error DFs for minus and plus
SE_L1_minus <- std.error(noOut_TloopL1_minus)
SE_L1_plus <- std.error(noOut_TloopL1_plus)
#create one data frame that has medians and standard errors
L1_all_summary_table <- cbind.data.frame(medianL1_minus, SE_L1_minus, medianL1_plus, SE_L1_plus, medianL1_FI)
#the measurements should have 2 sig figs and the standards errors should have 1 sig fig
odd_cols <- c(1,3,5)
even_cols <- c(2,4)
L1_all_summary_table[,odd_cols] <- signif(L1_all_summary_table[,odd_cols],2)
L1_all_summary_table[,even_cols] <- signif(L1_all_summary_table[,even_cols],1)
#rename column names
colnames(L1_all_summary_table) <- c("-CNCbl", "-CNCbl_SE","+CNCbl", "+CNCbl_SE", "FI")
#write csv
write.csv(L1_all_summary_table,"~/Downloads/Tloop_CNCbl/output_files/L1_all_summary_file.csv", row.names = TRUE)
#remove the switches below the functional cutoff of 2.0
#find the FI median
L1_summary_table <- subset(L1_all_summary_table, FI>1.9)
#write csv
write.csv(L1_summary_table,"~/Downloads/Tloop_CNCbl/output_files//L1_summary_file.csv", row.names = TRUE)
#write excel
write.xlsx(L1_summary_table, '~/Downloads/Tloop_CNCbl/output_files/L1_summary_file.xlsx', rowNames = TRUE)
write.xlsx(L1_all_summary_table, '~/Downloads/Tloop_CNCbl/output_files/L1_all_summary_file.xlsx', rowNames = TRUE)
##LIBRARY 2
#median DFs
medianL2_FI<- as.data.frame(apply(noOut_TloopL2_FI, 2, median, na.rm=TRUE))
medianL2_minus<- as.data.frame(apply(noOut_TloopL2_minus, 2, median, na.rm=TRUE))
medianL2_plus<- as.data.frame(apply(noOut_TloopL2_plus, 2, median, na.rm=TRUE))
#standard error DFs for minus and plus
SE_L2_minus <- std.error(noOut_TloopL2_minus)
SE_L2_plus <- std.error(noOut_TloopL2_plus)
#create one data frame that has medians and standard errors
L2_all_summary_table <- cbind.data.frame(medianL2_minus, SE_L2_minus, medianL2_plus, SE_L2_plus, medianL2_FI)
#the measurements should have 2 sig figs and the standards errors should have 1 sig fig
odd_cols <- c(1,3,5)
even_cols <- c(2,4)
L2_all_summary_table[,odd_cols] <- signif(L2_all_summary_table[,odd_cols],2)
L2_all_summary_table[,even_cols] <- signif(L2_all_summary_table[,even_cols],1)
#rename column names
colnames(L2_all_summary_table) <- c("-CNCbl", "-CNCbl_SE","+CNCbl", "+CNCbl_SE", "FI")
#write csv
write.csv(L2_all_summary_table,"~/Downloads/Tloop_CNCbl/output_files/L2_all_summary_file.csv", row.names = TRUE)

#remove the switches below the functional cutoff of 2.0
#find the FI median
L2_summary_table <- subset(L2_all_summary_table, FI>1.9)
#write csv
write.csv(L2_summary_table,"~/Downloads/Tloop_CNCbl/output_files/L2_summary_file.csv", row.names = TRUE)
#write excel
write.xlsx(L2_summary_table, '~/Downloads/Tloop_CNCbl/output_files/L2_summary_file.xlsx', rowNames = TRUE)
write.xlsx(L2_all_summary_table, '~/Downloads/Tloop_CNCbl/output_files/L2__all_summary_file.xlsx', rowNames = TRUE)
##LIBRARY 3
#median DFs
medianL3_FI<- as.data.frame(apply(noOut_TloopL3_FI, 2, median, na.rm=TRUE))
medianL3_minus<- as.data.frame(apply(noOut_TloopL3_minus, 2, median, na.rm=TRUE))
medianL3_plus<- as.data.frame(apply(noOut_TloopL3_plus, 2, median, na.rm=TRUE))
#standard error DFs for minus and plus
SE_L3_minus <- std.error(noOut_TloopL3_minus)
SE_L3_plus <- std.error(noOut_TloopL3_plus)
#create one data frame that has medians and standard errors
L3_all_summary_table <- cbind.data.frame(medianL3_minus, SE_L3_minus, medianL3_plus, SE_L3_plus, medianL3_FI)
#the measurements should have 2 sig figs and the standards errors should have 1 sig fig
odd_cols <- c(1,3,5)
even_cols <- c(2,4)
L3_all_summary_table[,odd_cols] <- signif(L3_all_summary_table[,odd_cols],2)
L3_all_summary_table[,even_cols] <- signif(L3_all_summary_table[,even_cols],1)
#rename column names
colnames(L3_all_summary_table) <- c("-CNCbl", "-CNCbl_SE","+CNCbl", "+CNCbl_SE", "FI")
#write csv
write.csv(L3_all_summary_table,"~/Downloads/Tloop_CNCbl/output_files/L3_all_summary_file.csv", row.names = TRUE)

#remove the switches below the functional cutoff of 2.0
#find the FI median
L3_summary_table <- subset(L3_all_summary_table, FI>1.9)
#write csv
write.csv(L3_summary_table,"~/Downloads/Tloop_CNCbl/output_files/L3_summary_file.csv", row.names = TRUE)
#write excel
write.xlsx(L3_summary_table, '~/Downloads/Tloop_CNCbl/output_files/L3_summary_file.xlsx', rowNames = TRUE)
write.xlsx(L3_all_summary_table, '~/Downloads/Tloop_CNCbl/output_files/L3_all_summary_file.xlsx', rowNames = TRUE)
##LIBRARY 4
#median DFs
medianL4_FI<- as.data.frame(apply(noOut_TloopL4_FI, 2, median, na.rm=TRUE))
medianL4_minus<- as.data.frame(apply(noOut_TloopL4_minus, 2, median, na.rm=TRUE))
medianL4_plus<- as.data.frame(apply(noOut_TloopL4_plus, 2, median, na.rm=TRUE))
#standard error DFs for minus and plus
SE_L4_minus <- std.error(noOut_TloopL4_minus)
SE_L4_plus <- std.error(noOut_TloopL4_plus)
#create one data frame that has medians and standard errors
L4_all_summary_table <- cbind.data.frame(medianL4_minus, SE_L4_minus, medianL4_plus, SE_L4_plus, medianL4_FI)
#the measurements should have 2 sig figs and the standards errors should have 1 sig fig
odd_cols <- c(1,3,5)
even_cols <- c(2,4)
L4_all_summary_table[,odd_cols] <- signif(L4_all_summary_table[,odd_cols],2)
L4_all_summary_table[,even_cols] <- signif(L4_all_summary_table[,even_cols],1)
#rename column names
colnames(L4_all_summary_table) <- c("-CNCbl", "-CNCbl_SE","+CNCbl", "+CNCbl_SE", "FI")
#write csv
write.csv(L4_all_summary_table,"~/Downloads/Tloop_CNCbl/output_files/L4_all_summary_file.csv", row.names = TRUE)

#remove the switches below the functional cutoff of 2.0
#find the FI median
L4_summary_table <- subset(L4_all_summary_table, FI>1.9)
#write csv
write.csv(L4_summary_table,"~/Downloads/Tloop_CNCbl/output_files/L4_summary_file.csv", row.names = TRUE)
#write excel
write.xlsx(L4_summary_table, '~/Downloads/Tloop_CNCbl/output_files/L4_summary_file.xlsx', rowNames = TRUE)
write.xlsx(L4_all_summary_table, '~/Downloads/Tloop_CNCbl/output_files/L4_all_summary_file.xlsx', rowNames = TRUE)
##LIBRARY 5
#median DFs
medianL5_FI<- as.data.frame(apply(noOut_TloopL5_FI, 2, median, na.rm=TRUE))
medianL5_minus<- as.data.frame(apply(noOut_TloopL5_minus, 2, median, na.rm=TRUE))
medianL5_plus<- as.data.frame(apply(noOut_TloopL5_plus, 2, median, na.rm=TRUE))
#standard error DFs for minus and plus
SE_L5_minus <- std.error(noOut_TloopL5_minus)
SE_L5_plus <- std.error(noOut_TloopL5_plus)
#create one data frame that has medians and standard errors
L5_all_summary_table <- cbind.data.frame(medianL5_minus, SE_L5_minus, medianL5_plus, SE_L5_plus, medianL5_FI)
#the measurements should have 2 sig figs and the standards errors should have 1 sig fig
odd_cols <- c(1,3,5)
even_cols <- c(2,4)
L5_all_summary_table[,odd_cols] <- signif(L5_all_summary_table[,odd_cols],2)
L5_all_summary_table[,even_cols] <- signif(L5_all_summary_table[,even_cols],1)
#rename column names
colnames(L5_all_summary_table) <- c("-CNCbl", "-CNCbl_SE","+CNCbl", "+CNCbl_SE", "FI")
#write csv
write.csv(L5_all_summary_table,"~/Downloads/Tloop_CNCbl/output_files/L5_all_summary_file.csv", row.names = TRUE)

#remove the switches below the functional cutoff of 2.0
#find the FI median
L5_summary_table <- subset(L5_all_summary_table, FI>1.9)
#write csv
write.csv(L5_summary_table,"~/Downloads/Tloop_CNCbl/output_files/L5_summary_file.csv", row.names = TRUE)
#write excel
write.xlsx(L5_summary_table, '~/Downloads/Tloop_CNCbl/output_files/L5_summary_file.xlsx', rowNames = TRUE)
write.xlsx(L5_all_summary_table, '~/Downloads/Tloop_CNCbl/output_files/L5__all_summary_file.xlsx', rowNames = TRUE)
##TOP SWITCHERS
#median DFs
medianTop10_FI<- as.data.frame(apply(noOut_TloopTop10_FI, 2, median, na.rm=TRUE))
medianTop10_minus<- as.data.frame(apply(noOut_TloopTop10_minus, 2, median, na.rm=TRUE))
medianTop10_plus<- as.data.frame(apply(noOut_TloopTop10_plus, 2, median, na.rm=TRUE))
#standard error DFs for minus and plus
SE_Top10_minus <- std.error(noOut_TloopTop10_minus)
SE_Top10_plus <- std.error(noOut_TloopTop10_plus)
#create one data frame that has medians and standard errors
Top10_all_summary_table <- cbind.data.frame(medianTop10_minus, SE_Top10_minus, medianTop10_plus, SE_Top10_plus, medianTop10_FI)
#the measurements should have 2 sig figs and the standards errors should have 1 sig fig
odd_cols <- c(1,3,5)
even_cols <- c(2,4)
Top10_all_summary_table[,odd_cols] <- signif(Top10_all_summary_table[,odd_cols],2)
Top10_all_summary_table[,even_cols] <- signif(Top10_all_summary_table[,even_cols],1)
#rename column names
colnames(Top10_all_summary_table) <- c("-CNCbl", "-CNCbl_SE","+CNCbl", "+CNCbl_SE", "FI")
#write csv
write.csv(Top10_all_summary_table,"~/Downloads/Tloop_CNCbl/output_files/Top10_all_summary_file.csv", row.names = TRUE)
#remove the switches below the functional cutoff of 2.0
#find the FI median
Top10_summary_table <- subset(Top10_all_summary_table, FI>1.9)
#write csv
write.csv(Top10_summary_table,"~/Downloads/Tloop_CNCbl/output_files/Top10_summary_file.csv", row.names = TRUE)
#write excel
write.xlsx(Top10_summary_table, '~/Downloads/Tloop_CNCbl/output_files/Top10_summary_file.xlsx', rowNames = TRUE)
write.xlsx(Top10_all_summary_table, '~/Downloads/Tloop_CNCbl/output_files/Top10__all_summary_file.xlsx', rowNames = TRUE)
```


